One-armed spirals in locally isothermal, radially structured self-gravitating discs

### 2015-01-15 paper/response.txt

The calculations for this section has been moved in
Appendix C. Here, I state the finding that co-rotation and Lindblad
resonances are not important for the one-armed spiral pattern in our
disc models. This supports the notion that the instability is not
gravitational.

Section 7.3:

Here I discuss possible relevance of the results to protoplanetary
discs.

I state that the destabilising mechanism --- the background
torque --- exists in the presence of a fixed temperature
gradient. (It only vanishes identically for strictly 
adiabatic/barotropic discs, see Appendix A.) This in turn requires the system t
o return to its initial temperature profile quickly when perturbed.

@@

The other requirement, more specific to trapping one-armed
spirals, is the radial disc structure. Our disc models can trap
one-armed modes because of the rising Q-profile towards the inner disc
which, in turn, is a result of the adopted surface density bump.

A surface density bump may represent mass built-up at some radial
location in the disc. Protoplanetary disc dead zones are an obvious
example. However, as the reviewer points out, dead zones may not
support the background torque because it cools too slowly.

Actually, recent simulations performed by Bae et
al. (2014) show that dead zones develop large-scale gravitational
instabilities, as opposed to high-m, small-scale local
gravito-turbulence.

I suggest planetary gaps as another possibility. For example, if a
giant planet opens a gap in a constant-Q disc, then the trapping
condition Eq. 49, may give rise to two Q-barriers at the outer gap
edge. Note that this does not necessarily require strong self-gravity.

### 2015-01-15 paper/response.txt

The numerical setup was chosen to effectively demonstrate the
destabilising effect of the background torque (Section 5.2 and
5.3). However, its occurence is not limited to the particular setup
presented in Section 5 and 6.

I have performed additional simulations with 
higher Q and simulations with only an inner bump edge (by setting a
large R2, see Eq. 33). These simulations still developed the one-armed
spiral, but require longer runs so high-resolution/3D is not
feasible. A full parameter survey is beyond the scope of this work,
but is the subject of a follow-up study.

***************************************************************************
It's my view that if you used a realistic equation of state that
included heating and cooling, the high-m mode instability would either
persist (as suggested by Lodato & Rice 2004) or the system would
simply heat up until it became stable against the growth of any
perturbation.  The disc mass is also quite low (0.086M*) and if one
were to estimate the disc temperature using some kind of realistic
assumption about the central star mass and disc outer radius, my
suspicion is that to have the Q profile used, the outer radius would
either have to be very small (a very compact disc) or the disc
temperature would be unrealistically low.  
***************************************************************************
The high-m perturbations do in fact persist in the reference run 
(Fig. 2), but it is overshadowed by the one-armed instability.

Lodato & Rice (2004) cool their discs to zero temperature
which is balanced against shock-heating (as in many similar
studies). The systems we consider are closer to irradiated discs (Rice
et al. 2011) in which the temperature is set externally 
(e.g. Stamatellos &  Whitworth 2008, Boley 2009). I note that 
such global simulations including heating (irradiation) and cooling 
may not have been run long enough to see low-frequency modes 
which grow slowly, and thus appear stable.

### 2015-01-15 paper/discussion.tex
nearly done with first revision

  %chiang and goldreich 
  The background torque may thus exist in the outer
  parts of protoplanetary discs that are irradiated, such
  that the disc temperature is set externally with a non-zero radial
  gradient. Of course, if external irradiation sets a strictly
  isothermal outer disc \citep[e.g.][]{boley09}, then the background
  torque vanishes.

>>>>

  %chiang and goldreich 
  The background torque may thus exist in the outer
  parts of protoplanetary discs that are irradiated, such
  that the disc temperature is set externally with a non-zero radial
  gradient \citep[e.g.][]{stamatellos08}. Of course, if external irradiation sets a strictly
  isothermal outer disc \citep[e.g.][]{boley09}, then the background
  torque vanishes.
